Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-08-05 idp/documentation/paper.tex
removed comments
fixed some grammar mistakes

			At this point, it seems like the two ways in which curve design software can affect the design process are through the choice of language and fairness measure. Ideally, the language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently. The fairness should capture the intuitive notion of curve smoothness sufficiently well. These goals are competing with ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures. From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.

@@

			% overview of bézier spline design tools
			% applying curve design model to bézier spline design tools
			% analysis of bézier spline design tools in terms of criteria
			%   works well: read, speak, derive
			%   has issues
			%     fairness
			%       curvature continuity is not ensured
			%       ensuring curvature continuity removes some degree of freedom, curves need more nodes
			%     expressiveness
			%       curves that are best described in terms of curvature
			%         need many nodes
			%           compromises smoothness
			%           makes the curve hard to change
			%           no curvature continuity guaranteed at nodes
			%         are difficult to specify, mistakes are easy
			\subsubsection{Common Design Tools for Bézier Splines}

@@

				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guarantee curvature continuity. It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it very difficult to create smooth curves if the design software does not support the user in this regard. The problem is aggravated by the fact that curvature incontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes. In case the software allows the user to request curvature-continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result. Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with language expressiveness.

				The language used to specify Bézier splines in common design tools is lacking in expressiveness. Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature). Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment. There is also no way to directly specify linearly changing curvature. This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result. Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created, respectively moved in order to change the overall curve. It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes. Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous. In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points, making it easy to make small mistakes.

@@

				Examples 2 and 3 show curves that could be described more efficiently in terms of curvature. Since there is no way to directly specify curvature using Bézier splines, a considerable number of nodes is needed to get close to the desired curve. It may be observed that the spiral in example 2 is not perfect. Creating perfect spirals using Bézier splines is very tedious and would require even more nodes (at least one for every 90\textdegree of winding angle covered). It could be speculated on whether the artist would have opted to use a perfect spiral in this example if they had had the tools efficiently do so.

>>>>

			At this point, it seems like the two ways in which curve design software can affect the design process are through the choice of language and fairness measure. Ideally, the language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently. The fairness should capture the intuitive notion of curve smoothness sufficiently well. These goals are competing with the software's ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures. From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.

@@

			\subsubsection{Common Design Tools for Bézier Splines}

@@

				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing curvature continuity. It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard. The problem is aggravated by the fact that curvature incontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes. In case the software allows the user to request curvature-continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result. Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with language expressiveness.

				The language used to specify Bézier splines in common design tools is lacking in expressiveness. Specifically, describing curves in terms of their curvature is difficult since curvature may only be specified indirectly through the magnitude of the velocity values (higher velocity resulting in lower curvature). Unfortunately, this is also very dependent on the point and velocity coefficients at the other end of the spline segment. There is also no way to directly specify linearly changing curvature. This makes specifying curvature very tedious, and in case the winding angle of the curve segment is large enough, actually impossible, forcing the designer to use additional nodes in order to get the desired result. Using an excessive number of nodes makes it more tedious and time-consuming for the designer to create and/or modify such curves, since many nodes and handles have to be created, or moved in order to change the overall curve. It also sacrifices smoothness, since it's easy to accidently introduce subtle bumps in the curve when using a large number of nodes. Another issue is that without software support for curvature continuity, using more nodes usually also introduces more points where curvature is not continuous. In cases where the curvature is an important property of the curve (for example the constant curvature of circle arcs, or the linearly changing curvature of spirals), it may be very difficult to specify these properties indirectly using the coefficient points, making it easy to make small mistakes.

@@

				Examples 2 and 3 show curves that could be described more efficiently in terms of curvature. Since there is no way to directly specify curvature using Bézier splines, a considerable number of nodes is needed to get close to the desired curve. It may be observed that the spiral in example 2 is not perfect. Creating perfect spirals using Bézier splines is very tedious and would require even more nodes (at least one for every 90\textdegree of winding angle covered). It could be speculated on whether the artist would have opted to use a perfect spiral in this example if they had had the tools to efficiently do so.
